---
title: "KKP - KKCS"
author: "Workshop"
format: html

editor: visual
embed-resources: true  # bundle everything into one HTML file

# ──────────────────────────────────────────────────────────────
# Optional settings — uncomment to try:

# format:
#   html:
#     theme:
#       dark: darkly
#       light: cosmo # dark and ligth mode

#
# format:
#   html:
#     toc: true              # add a table of contents
#     number-sections: true  # number the headings
#     code-fold: show        # make code blocks collapsible
#   docx: default            # also produce a Word document
#
# execute:
#   echo: true               # show code in the output (default)
#   eval: true               # run the code (default)
#   warning: false           # hide warning messages
#   message: false           # hide package loading messages
# ──────────────────────────────────────────────────────────────
---

## Introduction

This document explores the **Oats** dataset — a classic agricultural experiment on oat yield. We will create tables, figures, and a simple model to demonstrate how *Quarto* integrates code, text, and output in a single file.

For more examples of what Quarto can produce, visit the [Quarto Gallery](https://quarto.org/docs/gallery/).

## Load Data

```{r}
#| label: load-data

data("Oats", package = "nlme")
Oats <- as.data.frame(Oats)
str(Oats)
head(Oats)
```

This dataset comes from a split-plot experiment on oat yield (Yates, 1935). It contains **3 varieties** (Golden Rain, Marvellous, Victory), **4 nitrogen levels** (0, 0.2, 0.4, 0.6 cwt/acre), and **6 blocks** — 72 observations in total.


## Inline Code

The dataset contains **`r nrow(Oats)`** observations. The overall mean yield was **`r round(mean(Oats$yield), 1)`** bushels/acre, ranging from `r min(Oats$yield)` to `r max(Oats$yield)`.

## Visualization

### Base R

```{r}
#| label: fig-yield-base
#| fig-cap: "Oat yield by nitrogen application rate (base R)"
#| fig-width: 8
#| fig-height: 5
#| eval: true

boxplot(yield ~ nitro, data = Oats,
        xlab = "Nitrogen (cwt/acre)", ylab = "Yield (bu/acre)",
        main = "Oat Yield by Nitrogen Level",
        col = c("#E8B4B8", "#B8D4E3", "#93C5B8", "#F5D6A8"))
```

### ggplot2

```{r}
#| label: fig-yield
#| fig-cap: "Oat yield by nitrogen application rate"
#| fig-width: 8
#| fig-height: 5

ggplot2::ggplot(Oats, ggplot2::aes(x = factor(nitro), y = yield,
                                    fill = factor(nitro))) +
  ggplot2::geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  ggplot2::geom_jitter(width = 0.1, alpha = 0.3) +
  ggplot2::labs(x = "Nitrogen (cwt/acre)", y = "Yield (bu/acre)") +
  ggplot2::theme_minimal(base_size = 14)
```

As shown in @fig-yield, yield increases with nitrogen application.

```{r}
#| label: fig-scatter
#| fig-cap: "Nitrogen response by oat variety"
#| fig-width: 9
#| fig-height: 5

ggplot2::ggplot(Oats, ggplot2::aes(x = nitro, y = yield, color = Variety)) +
  ggplot2::geom_point(size = 2, alpha = 0.6) +
  ggplot2::geom_smooth(method = "lm", se = F) +
  ggplot2::labs(x = "Nitrogen (cwt/acre)", y = "Yield (bu/acre)") +
  ggplot2::theme_minimal(base_size = 14)
```

@fig-scatter shows that all three varieties respond positively to nitrogen, with similar slopes.

## Simple Model

```{r}
#| label: fit-model

model <- lm(yield ~ nitro + Variety, data = Oats)
summary(model)
```

```{r}
#| label: tbl-model
#| tbl-cap: "Linear model coefficients"

summary(model)$coefficients |>
  knitr::kable(digits = 3)
```

The model (@tbl-model) suggests that each additional unit of nitrogen increases yield by approximately **`r round(coef(model)[2], 1)`** bushels/acre.

## Try Other Formats!

Try changing `format: html` to `format: docx` in the YAML header and render again. Same content, different output!
